Edge sign prediction in general graphs and connection with Correlation Clustering

### 2017-06-29 gtx_algorithm.tex
rewrite intro to galaxy tree

The \gtx{} algorithm takes as input an unweighted graph topology $G_0=(V_0, E_0)$ and produces a sequence of
graphs $\{G_t\}_{t=1}^T$ of decreasing size until each connected component of $G_0$ is reduced to a
single node. As we will prove later, after reaching this point, the algorithm has selected $|V_0| -
1$ edges that form a spanning tree of $G_0$. Intuitively, $G_{t+1}$ is obtained from $G_t$ by composing two
primitives so that we can informally write $G_{t+1} = \left(\collapseStar{} \circ
\extractStar{}\right)(G_t)$.
% [from a CMU NIPS submission http://www.stat.cmu.edu/~arinaldo/papers/mutualfriends.pdf] We
% consider Mutual-Friends to be *an algorithmic primitive*, by which we mean a kind of subroutine for
% a more complicated function that iterates Mutual-Friends, similarly to [5].

\extractStar{} partitions the graph $G_t$ into a set of stars and \collapseStar{} build the graph
made of those stars using the edges in $E_t$. We provide more details on those two operations in the
following, as well as their complexity analysis. Then we state formally the \gtx{} algorithm and prove
its termination and correctness.  Finally, we study its properties, such as the number of iterations
needed to finish and the stretch of the resulting tree. For simplicity and without loss of
generality, we assume that $G_0$ through $G_T$ consist of a single connected component.
\marginpars{Actually, we only need to assume that $G_0$ is connected because then we can prove that
the other subsequent $G_t$ will be as well.}

>>>>

To achieve our objective of building a spanning tree that minimizes the distances between all
connected pair of nodes in the original graph, we rely on a particular subgraph structure, namely
the star.  We therefore introduce two algorithmic primitives: \extractStar{}, which partition a
graph $G$ into a set of disjoint stars; and \collapseStar{}, which selects edges from $E$ to
assemble these stars into a new, smaller graph. Given a graph topology $G_0=(V_0, E_0)$ and assuming
for simplicity that $G_0$ consists of a single connected component,\footnote{For we can otherwise
run our algorithm in parallel on each connected components of $G_0$.} the \gtx{} algorithm
repeatedly applies these two primitives to produce a sequence of graphs $\{G_t\}_{t=0}^K$ of
decreasing size, until $G_K$ is made of a single node. All the edges selected while reaching
this stage then form the spanning tree we were looking for.
% [from a CMU NIPS submission http://www.stat.cmu.edu/~arinaldo/papers/mutualfriends.pdf] We
% consider Mutual-Friends to be *an algorithmic primitive*, by which we mean a kind of subroutine for
% a more complicated function that iterates Mutual-Friends, similarly to [5].

In the following, we provide a more precise description of our two primitives and analyze their
complexity. Then we state formally the complete \gtx{} algorithm, prove its termination and
correctness, and show a detailed example of its execution.  Finally, we study its properties, such
as the number of iterations needed to finish and the stretch of the resulting tree.

### 2017-06-29 lowstretch_and_cc.tex
rewrite intro to galaxy tree

In this section, we are interested in active learning algorithms, that first query a subset
$\etrain$ of the edges, observe the signs in $\etest$ and use them to predict the remaining signs.
More precisely, we focus on an algorithm that query a spanning tree $T$ of $G$ and predict the sign
of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
flipped, the longer the path in $T$, the more likely its parity will be not be equal to the truth
$\yuv{}$. Therefore we would like each such path to be as short as possible and in the rest of this
section, we describe a way to build spanning tree tailored for this situation.

\subsection{\gtx{}: a simple low stretch tree construction}
\label{sub:gtx_a_simple_low_stretch_tree_construction}

>>>>

In this section, we are interested in active learning algorithms, that first query a subset
$\etrain$ of the edges, observe the signs in $\etrain$ and use them to predict the remaining signs.
More precisely, we focus on an algorithm that queries spanning tree $T$ of $G$ and predicts the sign
of an edge $(u,v) \in \etest = E \setminus E_T$ as the parity of $\pathtuv$. While this
can\Todo{state it formally then ;)} be stated more formally~\autocite[Section
4.1]{Cesa-Bianchi2012b}, it makes sense intuitively that since each sign has been potentially
flipped, the longer the path in $T$, the more likely its parity will be not be equal to the true
sign $\yuv{}$. Therefore we would like each such path to be as short as possible. In the following,
we describe a way to build spanning trees tailored for this situation.

\subsection{\gtx{}: a spanning tree designed for sign prediction}
\label{sub:gtx_a_simple_low_stretch_tree_construction}
